\huge The human visual system preserves the hierarchy of 2-dimensional pattern regularity

### 2021-05-18 Writeup/kohler_clarke_2021.tex
small edits

A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e. transformations that preserve distance \citep{RN1425}. The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation (CITATION?). In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotation symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains reflection (m) or glide reflection (g). For example, $P4$ contains 4 fold rotation, while $P2M$ contains 2 fold rotation and one reflection (see Figure XXX). Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition. Full details of the naming convention and examples of the wallpaper groups can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia}}.

In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}. The full list of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material. Subgroup relationships between wallpaper groups can be distinguished by their indices. The index of a subgroup relationship is the number of cosets, i.e. the number of times the subgroup is found in the supergroup \citep{RN1425}. As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure 1C)}. If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained. \textit{P2} is thus a subgroup of \textit{P6}, and \textit{P6} can be generated by combining \textit{P2} with rotations \{$0^{\circ}$, $120^{\circ}$, $240^{\circ}$\}. Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}. The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.

\begin{figure}[hptb]
\centering
\includegraphics[width=0.75\linewidth]{../figures/figure1.pdf}
\caption{\textcolor{ForestGreen}{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A). All other relationships can be infered by identifying the shortest path through the hierarchy, and multiplying the subgroup indices. For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$. We also show enlarged versions of some of the subgroup relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
\end{figure}

>>>>

A wallpaper group is a topologically discrete group of isometries of the Euclidean plane, i.e. transformations that preserve distance \citep{RN1425}. The wallpaper groups differ in the number and kind of these transformations \textcolor{ForestGreen}{and we can uniquely refer to different groups using crystallographic notation. In brief, most groups are notated by $PXZ$, where $X\in\{1,2,3,4,6\}$ indicates the highest order of rotation symmetry and $Z\in\{m, g\}$ indicates whether the pattern contains reflection (m) or glide reflection (g). For example, $P4$ contains 4 fold rotation, while $P4MM$ contains 4 fold rotation and two mirror reflections. By convention, many of the groups are given shortened names: for example, $P4MM$ is usually referred to as $P4M$, as the second reflection can be deduced from the presence of a 4 fold rotation alongside a reflection.  (see Figures \ref{fig:subgroups} and \ref{fig:example_stimuli} for examples). Two of the groups start with a $C$ rather than a $P$, ($CM$ and $CMM$) which indicates that the symmetries are specified relative to a cell that itself contains repetition. Full details of the naming convention and examples of the wallpaper groups can be found on \href{https://en.wikipedia.org/wiki/Wallpaper_group}{wikipedia}}.

In mathematical group theory, when the elements of one group is completely contained in another, the inner group is called a subgroup of the outer group \citep{RN1425}. The full list of subgroup relationships is listed in Section 1.4.2 of the Supplementary Material. Subgroup relationships between wallpaper groups can be distinguished by their indices. The index of a subgroup relationship is the number of cosets, i.e. the number of times the subgroup is found in the supergroup \citep{RN1425}. As an example, let us consider groups \textit{P2} and \textit{P6} \textcolor{ForestGreen}{(see Figure 1C)}. If we ignore the translations in two directions that both groups share, group \textit{P6} consists of the set of rotations \{$0^{\circ}$, $60^{\circ}$, $120^{\circ}$, $180^{\circ}$, $240^{\circ}$, $300^{\circ}$\}, in which \textit{P2} \{$0^{\circ}$, $180^{\circ}$\} is contained. \textit{P2} is thus a subgroup of \textit{P6}, and \textit{P6} can be generated by combining \textit{P2} with rotations \{$0^{\circ}$, $120^{\circ}$, $240^{\circ}$\}. Because \textit{P2} is repeated three times in \textit{P6}, \textit{P2} is a subgroup of \textit{P6} with index 3 \citep{RN1425}. \textcolor{ForestGreen}{A more complex example are the subgroups of \textit{PMM}, which is made up of two reflections, and a rotation of $180^{\circ}$. As such, $P2$ and $PM$ are both subgroups of $PMM$. (Illustrative examples are given in Figure \ref{fig:subgroups})} The 17 wallpaper groups thus obey a hierarchy of complexity where simpler groups are subgroups of more complex ones \citep{RN1711}.

\begin{figure}[hptb]
\centering
\includegraphics[width=0.75\linewidth]{../figures/figure1.pdf}
\caption{\textcolor{ForestGreen}{Subgroup relationships with indices 2 (solid lines) and 3 (dashed line) are shown in (A). All other relationships can be infered by identifying the shortest path through the hierarchy, and multiplying the subgroup indices. For example, \textit{P1} is related to \textit{P6} through \textit{P6}\textrightarrow\textit{P3} (index 2) and \textit{P3}\textrightarrow\textit{P1} (index 3) so \textit{P1} is also a subgroup of \textit{P6} with index $3 \times 2 = 6$. We also show enlarged versions of some of the subgroup relationships involving \textit{PMM} (B, shown in blue) and \textit{P6} (C, shown in red) and highlight the symmetries within the subgroups to emphasize how the supergroup can be generated by adding additional transformations to the subgroup.}}
\label{fig:subgroups}
\end{figure}
